---
title: "Getting Started with Tidymodels"
author: 'Luca Baggi'
output: html_notebook
---

# Load the libraries:

```{r}
# Main framework
library(tidymodels)

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
```

And set the seed;

```{r}
set.seed(42)
```

# Load the data

```{r}
urchins <- read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  # Change the columns names to be a little more verbose
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  # Factors are very helpful for modeling, so we convert one column
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))
```

Let's use `prop.table()` to show the ratios of the observations per `food_regime`:

```{r}
table(urchins$food_regime) %>% prop.table()
```

This is lucky, but it happens because the data has been purposefully created.

# Visualisations

```{r}
urchins %>%
  ggplot() +
  # colorise by food regime
  aes(initial_volume, width, col = food_regime) +
  geom_point() +
  # and add a linear regression
  geom_smooth(method = 'lm')
```

We leave the confidence intervals (CI) as they indicate where the estimate is distorted by a lack of data in a specific region. Still, it is clear that there might be a pattern for classification!


# Model fitting
No data pre-processing is done in this tutorial; we will predict a new value outside of this dataset.

We define our linear model:

```{r}
lm_urchins <- linear_reg() %>%
  set_engine('lm') %>%
  fit(width ~ initial_volume * food_regime, data = urchins)
```

And then visualise it in the `tidy` format:

```{r}
lm_urchins %>% tidy()
```

Also, note that the regression automatically adds the base terms and plots the coefficients for each level interaction.

## Better visualise the outcomes of the regression

This step is needed to generate a dot-and-whisker plot with `dwplot`. The vanilla command prints red elements. We can tweak it to make them a different colour. However, we can add a vertical line to indicate the zero level.

```{r}
lm_urchins %>% tidy() %>%
  dwplot(
    vline = geom_vline(xintercept = 0, colour = 'grey', lintype = 2)
  )
```

This signals that the only statistically significant variable is the initial volume and the interaction with the `low` food regime: all of the others cross the 0 boundary, so they are not different from zero at the 95% significance level.

# Model predictions

Let's create some new points:

```{r}
# expand.grid() is a base function that creates a dataframe out of the inputs' combinations.
new_points <- expand.grid(initial_volume = 20,
                          food_regime = c('Initial', 'Low', 'High'))
```

And see what comes out (a tibble):

```{r}
new_points
```

To make a prediction, let's get back to our model, `lm_urchins`, and apply it to the predictions with the `predict()` command. The command returns a tibble. **It has the same size as the `new_points` tibble, so we can bind the new columns to it!**.

```{r}
new_points <- lm_urchins %>%
  predict(new_data = new_points) %>%
  bind_cols(new_points)
```

We can also use the model to compute the predicted confidence intervals and bind them to the tibble of the predicted results!

```{r}
new_points <- lm_urchins %>%
  predict(new_data = new_points,
          type = 'conf_int') %>%
  bind_cols(new_points)
```

```{r}
new_points
```


And we can plot this with the `dwplot` package:

```{r}
new_points %>%
  ggplot() +
  aes(x = food_regime) +
  geom_point(aes(y = .pred)) +
  geom_errorbar(aes(ymin = .pred_lower,
                    ymax = .pred_upper),
                width = 0.2) +
  labs(y = 'urchins size')
```


# Using a different engine

We can use `stan` to create a bayesian linear model. We need to specify the `prior` distribution of the parameter (recall, this is the main difference between frequentist and bayesian analysis). We are given a Cauchy distribution, which is simply a t with one degree of freedom.

```{r}
# library(rstanarm) # error in installation, lol
```


```{r}
# param_distribution = rstanarm::student_t(df = 1)
```

